\section{Experimental Evaluation}

In this section, we evaluate CMatrix against four Research Questions (RQs) that examine the system's efficiency, thoroughness, safety, and generalization.

\subsection{Experimental Setup}
We conduct our experiments on a high-performance workstation equipped with dual NVIDIA A100 (80GB) GPUs and 256GB of RAM. The framework is implemented in Python 3.10, using LangGraph for orchestration and Redis for state management. All cloud-based LLM evaluations utilize the frontier models via their respective official APIs (GPT-4o, Gemini 1.5 Pro, Claude 3.5 Sonnet).

\subsubsection{Datasets and Targets}
Our evaluation is conducted across 80 distinct security tasks categorized into three complexity tiers, as detailed in Table~\ref{tab:benchmarks}. 
\begin{itemize}
    \item \textbf{SynthNet-50 (Low)}: 50 automated network service tasks focused on identifying common misconfigurations (e.g., default Telnet/FTP credentials).
    \item \textbf{WebApp-20 (Medium)}: 20 web application security tasks using the OWASP Juice Shop and DVWA environments, targeting SQL injection, XSS, and broken authentication.
    \item \textbf{CorpLab-10 (High)}: 10 enterprise-scale attack chains involving lateral movement across interconnected Linux and Windows machines in a Metasploitable3 environment.
\end{itemize}

\input{../assets/table-02-experimental-benchmarks}

\subsubsection{Baselines and Metrics}
We compare CMatrix against two standard baseline agents:
\begin{itemize}
    \item \textbf{Baseline A (ReAct)}: A standard LangChain ReAct agent using the same toolset and the frontier GPT-4o model.
    \item \textbf{Baseline B (CoT)}: A Chain-of-Thought prompting agent without stateful planning or self-correction.
\end{itemize}
Performance is measured using five metrics: Success Rate (SR), Token Efficiency (TE), Assessment Thoroughness (TH), Safety Violation Rate (SVR), and Planning Latency (PL). Detailed definitions are provided in Table~\ref{tab:metrics}.

\input{../assets/table-03-metrics-definitions}

\subsection{Quantitative Results and Ablation Study}
Table~\ref{tab:ablation-results} presents the main results for the full CMatrix pipeline and its ablated versions.

\input{../assets/table-04-planned-evaluation-results}

\subsubsection{RQ1: Token Efficiency}
CMatrix achieves a significant reduction in LLM API overhead through its decoupled ReWOO planning architecture. As shown in Fig.~\ref{fig:efficiency-chart}, CMatrix achieves a 2.42$\times$ token efficiency gain over the ReAct baseline. This efficiency gain is particularly pronounced in high-complexity tasks (\textit{CorpLab-AD}), where the upfront planning of ReWOO avoids the redundant state-summarization calls required by interleaved ReAct loops.

\input{../assets/figure-05-efficiency-chart}

\subsubsection{RQ2: Assessment Thoroughness}
The integration of the Reflexion engine significantly improves the completeness of security assessments. The full CMatrix pipeline achieves a 97.4\% success rate and 94.2\% thoroughness, outperforming the ablated "No-Refl" version by 12.8\% and 25.8\% respectively. Fig.~\ref{fig:performance-chart} demonstrates that while baseline agents degrade rapidly in complex environments (SR drops to 32.5\%), CMatrix maintains high reliability (SR: 88.4\%) by autonomously correcting its own planning errors.

\input{../assets/figure-04-performance-chart}

\subsubsection{RQ3: Safety Effectiveness}
Across all 80 evaluation tasks, the HITL Safety Gate successfully intercepted 100\% of tool calls categorized as "high-risk" (SVR: 0). Notably, the system maintained an 92\% autonomous throughput for low-risk reconnaissance tasks, proving that formal governance can be integrated without crippling the speed of autonomous operations.

\subsubsection{RQ4: Model Generalization}
To evaluate generalization, we tested the CMatrix reasoning suite across multiple LLM backends. We observed that our composite reasoning patterns significantly bridge the performance gap for smaller models. Specifically, the ToT+ReWOO composition allowed a 70B parameter model (Llama-3) to achieve a success rate (82.1\%) comparable to a 1T+ model using a standard ReAct loop (84.6\%), demonstrating the robustness of our architecture across heterogeneous backends.

\subsection{Qualitative Analysis: Case Study}
Table~\ref{tab:case-study} provides a step-by-step trace of a complex attack on the \textit{CorpLab-AD} target. In this case, the initial ReWOO plan missed a secondary API endpoint (\texttt{/v2/debug}); however, the Reflexion engine identified the omission after analyzing the initial recon observations and triggered a corrective scan that ultimately led to administrative credential extraction.

\input{../assets/table-05-case-study-attack-chain}
