\documentclass[conference,compsoc]{IEEEtran}
\IEEEoverridecommandlockouts
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{listings}
\usepackage{array}
\usepackage{etoolbox}
\usepackage{url}
\usepackage{stfloats}

% Final IEEE S&P Formatting Refinements
\renewcommand{\thesection}{\Roman{section}}
\renewcommand{\thesubsection}{\Alph{subsection}}
\renewcommand{\thesubsubsection}{\arabic{subsubsection})}
\renewcommand{\thetable}{\Roman{table}}

% Fix for Table label
\usepackage[labelsep=period]{caption}
\DeclareCaptionLabelFormat{tableprefix}{TABLE #2}
\captionsetup[table]{labelformat=tableprefix, labelsep=newline, justification=centering, font={sc,footnotesize}, skip=10pt}

% Fix for Figure prefix
\DeclareCaptionLabelFormat{figprefix}{Fig. #2}
\captionsetup[figure]{labelformat=figprefix, labelsep=period}

% Global sequential numbering
\newcounter{globalfig}
\AtBeginEnvironment{figure}{\refstepcounter{globalfig}}
\AtBeginEnvironment{lstlisting}{\refstepcounter{globalfig}}
\renewcommand{\lstlistingname}{Fig.}

% Page numbers
\pagestyle{plain}

\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

\lstdefinestyle{mystyle}{
    backgroundcolor=\color{backcolour},   
    commentstyle=\color{codegreen},
    keywordstyle=\color{magenta},
    numberstyle=\tiny\color{codegray},
    stringstyle=\color{codepurple},
    basicstyle=\ttfamily\footnotesize,
    breakatwhitespace=false,         
    breaklines=true,                 
    captionpos=b,                    
    keepspaces=true,                 
    numbers=left,                    
    numbersep=5pt,                  
    showspaces=false,                
    showstringspaces=false,
    showtabs=false,                  
    tabsize=2,
    frame=single,
    rulecolor=\color{black}
}
\lstset{style=mystyle}

\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{colortbl}
\usetikzlibrary{shapes,arrows,positioning,fit,backgrounds,calc}
\pgfplotsset{compat=1.18}

\begin{document}

\title{LLMOrch-VAPT: A Resilient, Multi-Agent Orchestration Framework for Autonomous Vulnerability Assessment and Penetration Testing}

\author{
\IEEEauthorblockN{Nishan Paul\IEEEauthorrefmark{1}\IEEEauthorrefmark{2}, John Doe\IEEEauthorrefmark{1}, and Jane Smith\IEEEauthorrefmark{3}}
\IEEEauthorblockA{\IEEEauthorrefmark{1}\textit{Research \& Development, Joblio Security Labs}, Dhaka, Bangladesh}
\IEEEauthorblockA{\IEEEauthorrefmark{2}\textit{Department of CSE, University of Dhaka}, Dhaka, Bangladesh}
\IEEEauthorblockA{\IEEEauthorrefmark{3}\textit{Open Source Security Collective}, San Francisco, USA \\
\{nishan.paul, j.doe\}@joblio.ai, smith@agentic-vapt.ai}
}

\maketitle

\input{../sections/01-introduction.tex}
\input{../sections/02-related-work.tex}
\input{../sections/03-threat-model.tex}
\input{../sections/04-methodology.tex}
\input{../sections/05-evaluation.tex}
\input{../sections/06-discussion.tex}
\input{../sections/07-conclusion.tex}

\section*{Acknowledgment}
This research was supported by Joblio Security Labs and the Open Source Security Collective.

\bibliographystyle{IEEEtran}
\bibliography{references}

\appendix
\section{Extended Infrastructure and Data Management}
\subsection{Detailed Agent Message-Passing Protocols}
Each agent in \texttt{LLMOrch-VAPT} utilizes a standardized JSON-based message format to communicate with the Master Supervisor. The supervisor maintains a global priority queue for sub-tasks, ensuring that reconnaissance results from the Network Agent are immediately available to the Web Agent for deeper analysis.

\subsection{Risk Classification Taxonomy}
The safety governance module maintains a comprehensive registry of tool-risk mappings. For instance, any tool categorized under \texttt{DANGEROUS\_TOOLS} (e.g., \texttt{sqlmap} with \texttt{--dbs} flag) triggers an immediate checkpoint and approval request via the HITL gate.

\subsection{Vector Store and Memory Management}
We utilize Qdrant for vector storage and BGE-large for embeddings. This allows agents to correlate current findings with historical data and threat intelligence, improving the precision of their reasoning paths and enabling the Security-Semantic Caching (SSC) mechanism.

\subsection{Infrastructure Setup}
All experiments were conducted on a workstation with 2x NVIDIA A100 (80GB) GPUs, 256GB RAM, and 10Gbps dedicated fiber connection to cloud provider endpoints. Local evaluations used the Ollama framework for Llama-3 deployment.

\end{document}
}
