\section{Introduction}

The increasing complexity of modern cloud-native infrastructures has outpaced the capabilities of traditional static vulnerability scanners. While these tools excel at identifying known software flaws through signature matching, they are fundamentally limited in their ability to simulate "hands-on-keyboard" adversaries who chain minor misconfigurations into multi-stage attack vectors. The emergence of Large Language Models (LLMs) offers a potential solution, enabling autonomous agents to interpret unstructured tool outputs and generate adaptive strategies in real-time.

\subsection{The Challenge of Long-Horizon Security Reasoning}
Despite their potential, current LLM-driven security agents face critical operational hurdles when applied to enterprise-scale environments. First, standard interleaved reasoning patterns like ReAct suffer from \textit{reasoning drift} and high latency during long-horizon tasks, where each tool observation necessitates a full model re-invocation. Second, existing agents often lack a \textit{self-correction loop} that translates verbal reflection into actionable tool re-execution, leading to missed vulnerabilities in complex attack surfaces. Finally, the autonomous nature of these agents introduces significant \textit{operational risk}, necessitating a formal governance model that can intercept dangerous commands without disrupting the overall assessment flow.

\subsection{Our Contribution: Governed Autonomy in CMatrix}
To address these challenges, we introduce \textbf{CMatrix}, a stateful, multi-agent orchestration framework designed for resilient and efficient autonomous red teaming. Our framework moves beyond simple agentic loops by implementing a property we define as \textbf{Governed Autonomy}: an architectural paradigm where reasoning and strategic planning are fully autonomous, yet all network-side tool executions are subject to verifiable security policies and human authorization gates. 

The core contributions of this work are as follows:
\begin{itemize}
    \item \textbf{Composite Reasoning Architecture:} We present a framework that unifies Tree of Thoughts (ToT), ReWOO, and Reflexion into a sequential orchestration pipeline, enabling strategic lookahead and high-efficiency planning for security tasks.
    \item \textbf{Closed-Loop Executable Reflection:} We introduce a structured self-critique mechanism that translates natural language reflections into formal tool re-execution, ensuring thorough coverage of the target attack surface.
    \item \textbf{Risk-Aware HITL Safety Governance:} We integrate a formal risk taxonomy directly into the stateful orchestration loop, providing a verifiable trust boundary that intercepts high-risk operations for human authorization.
    \item \textbf{Empirical Evaluation:} We provide the first systematic study of advanced reasoning pattern composition in applied cybersecurity, demonstrating a 97.4\% success rate and a 2.42$\times$ reduction in token overhead across 80 representative security scenarios.
\end{itemize}

The remainder of this paper is organized as follows: Section II discusses related work. Section III defines our formal threat model. Section IV details the CMatrix architecture and methodology. Section V presents our experimental results, followed by a discussion in Section VI and conclusion in Section VII.
