\section{Background and Related Work}
\label{sec:background}

The evolution of autonomous penetration testing has progressed from simple rule-based scripts to sophisticated multi-agent reasoning systems. In this section, we contextualize CMatrix within the current state-of-the-art.

\subsection{LLM-Driven Offensive Security}
Early efforts in this domain, such as PentestGPT \cite{pentestgpt} and HackingBuddyGPT \cite{hackingbuddygpt}, demonstrated that frontier models like GPT-4 could reason about high-level penetration testing goals and generate corresponding tool commands. However, these systems were largely monolithic and suffered from context window limitations during long engagements. Recent work, such as AutoAttacker \cite{autoattacker} and Cybench \cite{cybench}, has focused on benchmarking these capabilities across Capture-The-Flag (CTF) environments, revealing significant gaps in privilege escalation and lateral movement.

\subsection{Multi-Agent Orchestration}
To overcome the limitations of monolithic planners, researchers have proposed multi-agent frameworks. CHECKMATE \cite{checkmate2025} and xOffense \cite{xoffense2025} represent the current SOTA, utilizing hierarchical supervisor patterns to delegate specialized tasks to sub-agents. While these frameworks improve task completion rates, they operate in an "unconstrained" mode, where the orchestrator has full authority to execute any tool without safety governance.

\subsection{The Safety and Persistence Gap}
Despite advancements in capability, two critical gaps remain: (1) \textbf{Safety Governance}: There is a lack of formal frameworks to prevent agents from causing unintended operational harm; and (2) \textbf{Persistence}: Existing agents are typically stateless, losing their entire reasoning context if the process is interrupted or if a human intervention is required. CMatrix is designed specifically to fill these gaps through governed statefulness.

\input{assets/related_work_table.tex}

\section{Threat Model and Safety Properties}
\label{sec:threat_model}

We define a formal framework for assessing the safety and integrity of autonomous red teaming agents.

\subsection{System Model}
Our system consists of a \textit{Master Orchestrator} $\mathcal{M}$, a set of \textit{Specialized Workers} $\mathcal{W}$, and a \textit{Human Operator} $\mathcal{H}$. The goal of $\mathcal{M}$ is to fulfill a security objective $\mathcal{G}$ by executing a sequence of actions $\mathcal{A} = \{a_1, a_2, \dots, a_n\}$.

\subsection{Adversary Models}
We consider two primary adversary models that threaten the safety of the red teaming engagement:
\begin{itemize}
    \item \textbf{Adversary $\alpha$ (Indirect Prompt Injection)}: An external attacker who places malicious instructions in the target environment (e.g., a web page or configuration file). When the agent scrapes this content, it may be coerced into executing destructive actions (e.g., \texttt{rm -rf /}).
    \item \textbf{Adversary $\beta$ (Execution Fault/Hallucination)}: A non-malicious but dangerous state where the LLM's reasoning engine hallucinates a destructive command as a valid tactical step (e.g., misinterpreting a "wipe disk" utility as a "log analyzer").
\end{itemize}

\subsection{Safety Invariants}
To defend against these adversaries, CMatrix enforces four core safety properties ($P$):
\begin{itemize}
    \item \textbf{$P_1$: Pattern Integrity (Auto-Reject)}. No tool invocation $\Theta$ shall execute if it matches a pre-defined prohibited pattern list $\mathcal{P}_{reject}$.
    \item \textbf{$P_2$: Semantic Gating (HITL)}. No high-risk tool invocation $\Theta$ (where $L_{risk} \geq \text{High}$) shall execute without explicit authorization from $\mathcal{H}$.
    \item \textbf{$P_3$: Audit Completeness}. Every transition in the execution graph, including all approval/rejection decisions, must be persisted to a durable, tamper-evident state-store.
    \item \textbf{$P_4$: Context Preservation}. Resumption of a workflow after a HITL interruption must maintain 100\% of the reasoning context and conversation history.
\end{itemize}
CMatrix is designed to satisfy $P_1$--$P_4$ while maintaining an operational success rate comparable to unconstrained agents.
